---
title: "Licht-sediment relatie in de Schelde"
output: html_notebook
---

```{r}
require(tidyverse)
require(skimr)
require(leaflet)
require(sf)
```

## Data extraction

The Scheldemonitor was queried for parameters related to light extinction and suspended sediment.



```{r leesDataUitScheldemonitor, cache=TRUE, include=FALSE}
url <- "http://geo.vliz.be/geoserver/wfs/ows?service=WFS&version=1.1.0&request=GetFeature&typeName=Dataportal%3Aabiotic_observations&resultType=results&viewParams=where%3Aobs.context+%26%26+ARRAY%5B1%5D+AND+standardparameterid+IN+%283240%5C%2C461%5C%2C1293%5C%2C495%5C%2C3289%5C%2C1223%29%3Bcontext%3A0001&propertyName=stationname%2Clongitude%2Clatitude%2Cdatetime%2Cdepth%2Cparametername%2Cvaluesign%2Cvalue%2Cdataprovider%2Cdatasettitle&outputFormat=csv"

# df <- read_csv(url, na = c("-"))
# 
# df %>% arrange(-value) %>% head(1) %>% select(value) %>% as.character()
#
# nog een keer
df <- read_csv(url, na = c("-", "999999999999"))

# check of hoogste waarde niet eigenlijk NA moet zijn
df %>% arrange(-value) %>% head(1) %>% select(value) %>% as.character()
# check for other indicators of NA
df %>% filter(grepl("999", as.character(value))) %>% select(value) %>% as.character()

```

## Parameters

```{r}
df %>% distinct(parametername, datasettitle)
```

### Spatial extent per dataset

```{r}

pal = colorFactor(topo.colors(10), domain = as.factor(df$dataprovider))
df %>% distinct(latitude, longitude, dataprovider) %>%
  mutate(dataprovider = as.factor(dataprovider)) %>%
  filter(latitude>0) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    stroke = F, fillOpacity = 1, fillColor =  ~pal(dataprovider),
    label = ~dataprovider) %>%
  addLegend("bottomright", pal = pal, values = ~dataprovider,
    # title = "dataprovider",
    opacity = 1)
```


## Find duplicates

Data may be reported twice to Scheldemonitor by different organizations. 

```{r}
# duplicate rows?
df %>% 
  drop_na(value) %>%
  duplicated() %>% which() %>% length()
# no


# duplicate measurements?
df %>% 
  drop_na(value) %>%
  select(latitude, longitude, depth, datetime, parametername, stationname, value)  %>%
  duplicated() %>% which() %>% length()
# yes. 1

```

```{r remove duplicate measurements}
duplicates <- df %>% 
  select(latitude, longitude, depth, datetime, parametername, stationname, value)  %>%
  duplicated() %>% which()
df_clean <- df[-duplicates,]

# There may be multiple values at same location and time
df_clean %>% 
  drop_na(value) %>%
  select(latitude, longitude, depth, datetime, parametername, stationname)  %>%
  duplicated() %>% which() %>% length()
# yes, 54

# See if we can average them
dupliSamples <- df_clean %>% 
  drop_na(value) %>%
  select(latitude, longitude, depth, datetime, parametername, stationname)  %>%
  duplicated() %>% which()

df_clean %>%
  group_by(latitude, longitude, depth, datetime, parametername, stationname) %>%
  summarize(n = n(), mean = mean(value, na.rm = T), sd = sd(value, na.rm = T), .groups = "drop") %>%
  arrange(-n) %>% head(54)

# Some sd values are very high compared to mean. Still, averaging is necessary to work with the data further

# redefine df_clean without any duplicate measurements

df_clean <- df[-duplicates,] %>%
  select(latitude, longitude, depth, datetime, parametername, stationname, dataprovider, value) %>%
  group_by(latitude, longitude, depth, datetime, parametername, stationname, dataprovider) %>%
  summarize(mean = mean(value, na.rm = T), .groups = "drop")

```


There are a few samples with very high sd as compared to the mean (n = 2). 

Without further analysis, it is proposed to work with mean values. 



## Find matching parameters

Perhaps first aggregate parameters that can be aggregated?

```{r aggregate parameters}

df_clean %>% distinct(parametername)

```


Next step harmonizes the data with respect to units. Also, NA and zero values are dropped.

```{r}
df_harmonized <- df_clean %>%
  mutate(newParam = case_when(
    parametername == "Zwevend stof in mg/l in oppervlaktewater" ~ "SPM",
    parametername == "Zwevend stof in g/l in water" ~ "SPM",
    parametername == "Lichtextinctie in /m in oppervlaktewater" ~ "E",
    grepl("Absorptie", parametername) ~ "E",
    parametername == "Doorzicht in dm in oppervlaktewater" ~ "Secchi"
  )) %>%
    mutate(newValue = case_when(
    parametername == "Zwevend stof in mg/l in oppervlaktewater" ~ mean,
    parametername == "Zwevend stof in g/l in water" ~ mean/1000,
    parametername == "Lichtextinctie in /m in oppervlaktewater" ~ mean,
    grepl("Absorptie", parametername) ~ mean,
    parametername == "Doorzicht in dm in oppervlaktewater" ~ mean/10
  )) %>%
    mutate(newUnit = case_when(
    parametername == "Zwevend stof in mg/l in oppervlaktewater" ~ "mg/l",
    parametername == "Zwevend stof in g/l in water" ~ "mg/l",
    parametername == "Lichtextinctie in /m in oppervlaktewater" ~ "/m",
    grepl("Absorptie", parametername) ~ "/m",
    parametername == "Doorzicht in dm in oppervlaktewater" ~ "m"
  )) %>% 
  select(-parametername, -mean) %>%
  filter(newValue > 0 & newValue < 900) %>%
  drop_na(newValue)

```

Include organic carbon measurements

```{r}
df_C <- read_delim("data/orgC.csv", delim = ";")
df_harmonized <- df_harmonized %>% bind_rows(df_C)
```



Find parameters that were measured at the same time and location


```{r matchParameters}
# make cross-table
df_wide <- df_harmonized %>% 
  unite("paramUnit", newParam, newUnit) %>%
  spread(key = paramUnit, value = newValue)

View(df_wide)  

```


## Geoinformatie toevoegen

```{r}
df_clean %>% 
  distinct(latitude, longitude, stationname, dataprovider) %>%
  filter(latitude < 40)
```


Voor stations Schellebelle moeten nog coordinaten worden toegevoegd

## Additionele informatie

- watersysteem of OMES vak
- km van de monding?

```{r stationclassification}
stations <- df_clean %>% 
  distinct(latitude, longitude, stationname, dataprovider) %>%
  st_as_sf(coords = c("longitude", "latitude"),  crs = 4326)

OMESvakken <- st_read("data\\OMES_MOSES_gebiedsindeling_Evaluatiemethodiek_RD_new.shp") %>%
  st_transform(crs = 4326) %>% st_make_valid()

Oosterschelde <- st_read("data\\wl2018_OSv4.geojson")

OosterscheldeStations <- stations %>% st_intersection(Oosterschelde) %>% distinct(stationname) %>% st_drop_geometry() %>%
  mutate(watersysteem = "Oosterschelde")

scheldeStations <- stations %>% st_intersection(OMESvakken) %>% distinct(stationname, Omessegmen) %>% st_drop_geometry() %>%
  mutate(watersysteem = "Schelde")

labelledStations <-  OosterscheldeStations %>% 
  bind_rows(scheldeStations) %>% 
  right_join(stations)  %>% 
  select(-geometry) %>% distinct() %>%
  mutate(watersysteem = case_when(
    !is.na(watersysteem) ~ watersysteem,
    is.na(watersysteem) & dataprovider == "RWS - Rijkswaterstaat" ~ "Kustwater"
  ))



```

```{r combineStationMeasurements}
df_wide <- df_wide %>% left_join(labelledStations)

```



```{r saveHarmonizedData}

save(df_wide, file = "ScheldeSlib.Rdata")
```

